Binning Strategies and Related Loss for Large Data

### 2015-12-15 versionForJCGS/PaperJCGS.Rnw
Start to update Flight example section to discuss new 2014 flights data

\subsection{Big Data: Airline Departure Times}
The Federal Aviation Association (FAA) requires all airlines based in the United States to report   details for every single flight. These are published online by the Bureau of Transportation Services at \url{http://www.transtats.bts.gov/DataIndex.asp}. 
Every day there are about 16,000 flights across the United States adding up to almost 6 Million flights a year. Scheduled and actual departure times for all flights in 2011 make up --in uncompressed form-- a file of about 450 MB. A comparison of scheduled and actual departure times allows us an investigation of on-time performance of air carriers.

Figure~\ref{airline-scatter} shows two plots of the relationship between scheduled versus actual departure times. The plot on the left shows the scatterplot from a sample of one million of those records. Even while using alpha blending this results in a severely over-plotted graph. On the right is a minimally binned scatterplot of the reduced data from standard binning at 1-minute intervals. If the origin is offset by a half minute for both departures and arrivals, this binning does not have any spatial loss, as all observations will be centered within bins. The 1-minute standard binning also reduces the data to 176,384 reduced data triple, less than 3\% of the original data.

@@

\caption[Alpha-blended and minimally binned scatterplots for scheduled and actual departure times of airline flights.]{\label{airline-scatter}Scheduled and actual departure times of  flights across the United States in 2011. The plot on the left is based on a sample of the data, the plot on the right shows all flights. The large scale distributional patterns are visible in both plots, but the plot on the left misses some of the finer level details in scheduling that is visible in the plot on the right.}
\end{figure}

>>>>

\subsection{Big Data: Airline Departure Times}
The Federal Aviation Association (FAA) requires all airlines based in the United States to report   details for every single flight. These are published online by the Bureau of Transportation Services at \url{http://www.transtats.bts.gov/DataIndex.asp}. 
Every day there are about 16,000 flights across the United States which added up to over 5.8 million flights in 2014. The scheduled and actual departure times for flights in 2014 make up --in uncompressed comma separated form-- a file of about 53 MB. A comparison of scheduled and actual departure times allows us an investigation of on-time performance of air carriers.

Figure~\ref{airline-scatter} shows two plots of the relationship between scheduled versus actual departure times. The plot on the left shows the scatterplot from a sample of one million of those records. Even while using alpha blending this results in a severely over-plotted graph. On the right is a minimally binned scatterplot of the reduced data from standard binning at 1-minute intervals. If the origin is offset by a half minute for both departures and arrivals, this binning does not have any spatial loss, as all observations will be centered within bins. The 1-minute standard binning also reduces the data to around 190,000 reduced data triple, approximately 5\% of the original data size.

@@

\caption[Alpha-blended and minimally binned scatterplots for scheduled and actual departure times of airline flights.]{\label{airline-scatter}Scheduled and actual departure times of flights across the United States in 2014. The large scale distributional patterns are visible in both plots, but the plot on the left misses some of the finer level details in scheduling that is visible in the plot on the right.}
\end{figure}
